What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2024-08-22 manuscript.tex
little fixes to knit the pdf

\abstract{%
Web-based data collection allows researchers to recruit large and diverse samples with fewer resources than lab-based studies require. Recent innovations have expanded the set of methodolgies that are possible online, but ongoing work is needed to test the suitability of web-based tools for various research paradigms. Here, we focus on webcam-based eye-tracking; we tested whether the results of five different eye-tracking experiments in the cognitive psychology literature would replicate in a webcam-based format. Specifically, we carried out five experiments by integrating two javascript-based tools: js.psych and a modified version of Webgazer.js. In order to represent a wide range of applications of eye-tracking to cognitive psychology, we chose two psycholinguistic experiments, two memory experiments, and a decision-making experiment. These studies also varied in the type of eye-tracking display, including screens split into halves (Exps. 3 and 5) or quadrants (Exps. 2 \& 4), or composed scenes with regions of interest that varied in size (Exp. 1). Outcomes were mixed. The least successful replication attempt was Exp. 1; we did not obtain a condition effect in our remote sample (1a), nor in an in-lab follow-up (1b). However, the other four experiments were more successful, replicating a blank-screen effect (Exp. 2), a novelty preference (Exp. 3), a verb bias effect (Exp. 4), and a gaze-bias effect in decision-making (Exp. 5). These results suggest that webcam-based eye tracking can be used to detect a variety of cognitive phenomena, including those with sensitive time, although paradigms that require high spatial resolution should be adapted to coarser quadrant or split-half displays.
}

@@

During the calibration, dots appeared on the screen one at a time in
different locations, and the participants had to fixate them and click
on each one. Once they clicked on a dot, it would disappear and a new
one would appear in a different location on the screen. The locations of
calibration dots were specific to each experiment (details below) and
appeared in the areas of the screen where the visual stimuli would
appear during the main task in order to ensure that eye movements were
accurately recorded in the relevant regions of interest. After the
calibration was completed, the validation began. Participants were asked
to go through the same steps as the calibration, except that they only
fixated the dots as they appeared in different locations on the screen.
If accuracy on the validation was too low (fewer than 50\% of looks
landed within a 200 px radius of the validation points),
participants were given an opportunity to re-start the calibration and
validation steps. If the second attempt also lead to low validation
accuracy, participants were informed that they could not participate in
the study.

@@

\subsection{Data pre-processing}\label{data-pre-processing}

We used R (Version 4.4.1; R Core Team, 2021) and the R-packages \emph{afex} (Version 1.3.1; Singmann, Bolker, Westfall, Aust, \& Ben-Shachar, 2021), \emph{broom.mixed} (Version 0.2.9.5; Bolker \& Robinson, 2020), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, \& Müller, 2021), \emph{forcats} (Version 1.0.0; Wickham, 2021a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{jsonlite} (Version 1.8.8; Ooms, 2014), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lmerTest} (Version 3.1.3; Kuznetsova, Brockhoff, \& Christensen, 2017), \emph{Matrix} (Version 1.7.0; Bates \& Maechler, 2021), \emph{papaja} (Version 0.1.2; Aust \& Barth, 2020), \emph{readr} (Version 2.1.5; Wickham \& Hester, 2020), \emph{shiny} (Chang et al., 2021), \emph{stringr} (Version 1.5.1; Wickham, 2019), \emph{tidyr} (Version 1.3.1; Wickham, 2021b), and \emph{tinylabels} (Version 0.2.4; Barth, 2022) for all our analyses.

\section{Experiment 1}\label{experiment-1}

@@

We first collected data from participants via Prolific, then conducted a follow-up in-lab replication; both the remote and the in-lab replications used WebGazer and identical materials and procedures.

@@

\paragraph{Remote sample}\label{remote-sample}

@@

\paragraph{In-lab sample}\label{in-lab-sample}

Forty-nine participants completed the study in a lab setting. They were recruited via the college subject pool. Participants needed to be 18 years of age or older and native speakers of English. The study was approved by the Institutional Review Board at Boston College.

@@

\begin{verbatim}
## Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2
## 3.5.0.
## i Please use the `legend.position.inside` argument of `theme()` instead.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
## generated.
\end{verbatim}

@@

\subsubsection{Materials \& Design}\label{materials-design}

Each participant was presented with sixteen critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
sixteen fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.

\subsubsection{Data pre-processing and analysis}\label{data-pre-processing-and-analysis}

@@

\subsubsection{Remote sample}\label{remote-sample-1}

\paragraph{Minimal Exclusion}\label{minimal-exclusion}

The first set of analyses used minimal exclusion criteria. First, we eliminated participants with 0 percent of fixations in any ROIs. This resulted in the elimination of one participant. Second, we excluded participants with validation accuracy under 10 percent, resulting in an additional 5 excluded participants. The following analyses included 52 participants.

@@

For each sentence, the target time window began at the onset of the verb and ended 2000 milliseconds later. This window was then divided into 50-ms bins; for each participant and each trial, we recorded whether each object was fixated during the 50-ms bin. Collapsing over trials and participants, and averaging across distractors, we calculated the cumulative probability of fixation, shown in Figure \ref{fig:E1-spaghetti-fig}, Panel (b).

@@

In our first two analyses, we asked whether participants looked more to the target than to the distractor during the predictive time window, given that the verb is restricting. The first model tested whether there were more fixations to the target object than to the distractor in the time window before the onset of the target noun. We ran a regression model predicting the cumulative fixation probability in the last 50-ms bin before noun onset from the verb condition (restricting = 1 vs.~non-restricting = 0), object type (target = 1 vs.~distractor = 0), and their interaction, along with random effects for participants and images (with no covariance between random effects because the model cannot converge with full covariance matrix). There were no significant effects, although the critical interaction was in the expected direction (\emph{b} = 0.05, \emph{SE} = 0.03, \emph{p}=0.15).

@@

Altmann and Kamide tested a second model, aligning the predictive time window with the offset of the verb rather than the onset of the noun as above. When we did the same, we again saw that the critical interaction is not significant but numerically in the expected direction (\emph{b} = 0.05, \emph{SE} = 0.03, \emph{p}=0.20).

@@

Finally, we addressed whether participants looked to the target faster in the restrictive vs.~the non-restrictive condition, starting after the onset of the verb. On average, participants looked to the target 349 ms after the noun onset in the restrictive condition James, Minnihan, \& Watson (2023) and 452 ms after the noun onset in the non-restrictive condition James et al. (2023). Thus, first fixations were not only delayed relative to those in the previous studies compared here, but also showed a smaller difference between conditions.

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes. There were no significant effects; participants looked sooner at the target in the restrictive condition, while accounting for verb duration and its interaction with condition, but this was not a statistically significant effect (\emph{b} = -121.91, \emph{SE} = 90.57, \emph{p}=0.20).

\paragraph{Aggressive Exclusion}\label{aggressive-exclusion}

The second set of analyses used more aggressive exclusion criteria. First, we eliminated participants with 20 percent of their fixations or fewer landing in any ROIs. This resulted in the elimination of 15 participants. Second, we excluded participants with validation accuracy under 50 percent, which eliminated an additional 35 participants. The following analyses included 22 participants.

We tested the same three models under these more aggressive exclusion criteria. The first two models, comparing target and distractor fixations in the predictive window, produced very similar results; the critical interaction was not statistically significant (Pre-noun-onset window: \emph{b} = 0.07, \emph{SE} = 0.06, \emph{p}=0.23; Pre-verb-offset window: \emph{b} = 0.05, \emph{SE} = 0.05, \emph{p}=0.28). However, the final model, which tested the effect of verb condition on saccades to the target, yielded a statistically significant result, unlike in the previous set of analyses (\emph{b} = -193.35, \emph{SE} = 96.33, \emph{p}=0.05).

\paragraph{Calibration}\label{calibration}

@@

We tested whether a participant's calibration quality was correlated with their effect size. There were three effects of interest: the verb-by-object interaction in predicting fixation probabilities, both in the (1) pre-noun-onset and (2) pre-verb-offset windows (calculated as the difference in target-over-distractor preference between verb conditions), and (3) the effect of verb on the timing of the first target fixation (calculated as the difference in target latency between verb conditions). Across the three effects of interest, calibration quality was not significantly correlated (Effect 1: Pearson's \emph{r} = 0.03, \emph{p} = 0.83, Effect 2: Pearson's \emph{r} = -0.05, \emph{p} = 0.73, Effect 3: Pearson's \emph{r} = 0.04, \emph{p} = 0.78. In the remote sample, we saw that calibration quality was correlated with target advantage in the pre-noun window in the restricting condition alone. This was not the case in the in-lab sample (Pearson's \emph{r} = 0.21, \emph{p} = 0.14).

\subsubsection{In-lab sample}\label{in-lab-sample-1}

\paragraph{Minimal Exclusion}\label{minimal-exclusion-1}

As in the remote sample, we checked whether there were participants with 0 percent of fixations in any ROIs and there were none. We then excluded participants with validation accuracy under 10 percent, resulting in 2 excluded participants. The following analyses included 47 participants.

@@

For each sentence, the target time window began at the onset of the verb and ended 2000 milliseconds later. This window was then divided into 50-ms bins; for each participant and each trial, we recorded whether each object was fixated during the 50-ms bin. Collapsing over trials and participants, and averaging across distractors, we calculated the cumulative probability of fixation, shown in Figure \ref{fig:E1-spaghetti-fig}, Panel (c).

@@

Finally, we addressed whether participants looked to the target faster in the restrictive vs.~the non-restrictive condition, starting after the onset of the verb. On average, participants looked to the target 397 ms after the noun onset in the restrictive condition and 376 ms after the noun onset in the non-restrictive condition. As in the remote sample, the latencies are overall slower than in results published by Altmann \& Kamide (1999) and James et al.~(2023). Unlike in the remote sample, however, the difference is in the unexpected direction, such that participants looked to the target faster in the non-restrictive condition.

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes. There were no significant effects; results revealed that the paradoxical advantage in the non-restrictive condition was not statistically significant (\emph{b} = 21.70, \emph{SE} = 115.32, \emph{p}=0.85). Effects of verb duration and its interaction with condition were also not statistically significant (duration: \emph{b} = -0.58, \emph{SE} = 0.55, \emph{p}=0.30; interaction: \emph{b} = -0.49, \emph{SE} = 0.96, \emph{p}=0.61 )

\paragraph{Calibration}\label{calibration-1}

As before, participants' calibration quality was measured as the mean percentage of fixations that landed within 200 pixels of the calibration point. Calibration quality ranged from 5.13\% to 97.89\%.

We tested whether a participant's calibration quality was correlated with their effect size. Across the three condition effects of interest, calibration quality was not significantly correlated (pre-noun: Pearson's \emph{r} = -0.24, \emph{p} = 0.11; pre-verb-offset: Pearson's \emph{r} = -0.19, \emph{p} = 0.20; first fixation: Pearson's \emph{r} = -0.12, \emph{p} = 0.41. However, when the two interaction effects are calculated as the target advantage in the restricting condition only (i.e.~rather than a difference of differences), we see a significant correlation between target advantage and calibration quality in the wider pre-noun window (Pearson's \emph{r} = -0.16, \emph{p} = 0.29).

\paragraph{Aggressive Exclusion}\label{aggressive-exclusion-1}

The second set of analyses used more aggressive exclusion criteria. First, we eliminated participants with 20 percent of fixations or fewer landing in any ROIs. This resulted in the elimination of 7 participants. Second, we excluded participants with validation accuracy under 50 percent, which eliminated an additional 23 participants. The following analyses included 26 participants.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1-spaghetti-fig-1.pdf}
\caption{\label{fig:E1-spaghetti-fig}Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.}
\end{figure}

Overall, the results of Experiment 1 paint a sobering picture of web-based eye-tracking. In the remote sample, results were in the expected direction but effects were smaller and delayed relative to previous work and failed to reach statistical significance; the one exception was in comparing first target fixation latencies across verb conditions, but only when aggressive exclusion criteria were applied to participants (INSERT WHAT THOSE CRITERIA WERE). To test whether this could be explained by the variability in experimental settings across participants, we replicated our procedure in a lab setting. Surprisingly, the results in the in-lab study were less aligned with previous work; critical effects were not significant nor were they in the expected direction numerically. This suggests the problem was not with running the study online or due to working with a more diverse population.

Taken together, the instability of these effects might suggest that this paradigm is not well-suited for webcam-based eye-tracking. Notably, the ROIs were tightly drawn around the five to six objects in each scene (larger ROIs would have led to overlapping objects) and thus, analyses were unforgiving of inaccurate calibration. Further evidence comes from (Prystauka et al., 2023), who ran a modified version of the Altmann and Kamide study with only 4 objects, allowing larger, more distinct ROIs.

@@

60 participants were paid for their participation.
The sample size was motivated in part by budget constraints, but was
nonetheless 2.5x larger than the original sample size of 24).
Data from 1 participant were not properly
recorded due to unknown technical issues, so data from 59 participants
were included in all analyses to follow.

@@

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. During the encoding phase, participants saw a
grid indicating the four quadrants of the screen. Each quadrant
contained six images of items belonging to the same category (see Figure~\ref{fig:E2-example-trial}).
The four categories were humanoids, household objects, animals, and
methods of transportation.
Each of the four quadrants was presented one at a time. First, a list of
the items in the quadrant was shown, then the pictures of items were displayed in the quadrant.
For each item, participants used their arrow keys to indicate whether the object was facing left or right. After the participant identified the direction of each
item, they would have an additional 30 seconds to encode the name and
orientation of each item in the quadrant. Finally, after all four quadrants
were presented, participants were shown the full grid of
24 items and had 60 seconds to further encode the name and orientation
of each item.

During the recall phase, participants listened to statements and
responded by pressing the `F' key for false statements and `T' for true
ones. Each statement fell into either an interobject or intraobject
condition. Interobject statements were those that compared two different
items in the grid (e.g.~``The skeleton is to the left of the robot''),
while intraobject statements were those that asked about the orientation
of a single item (e.g.~``The bus is facing right''). There were 48 total
statements, with 24 interobject and 24 intraobject statements split
evenly among the four quadrants. While listening to these statements, in
the free-viewing block, participants saw a blank screen and were allowed
to freely gaze around the screen. During the fixed-viewing block,
participants were asked to fixate a small cross in the center of the
screen throughout the recall phase. In both cases, the mouse was
obscured from the screen. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first. Different images were used in each block.

After completing both encoding-recall blocks, participants were asked to
answer a few survey questions (such as whether they wore glasses or
encountered any distractions).

@@

\subsubsection{Replication}\label{replication}

\textbf{Eye-gaze}. Looks during the retrieval period were categorized as belonging to one of four quadrants based on the x,y coordinates. The critical quadrant was the one in which the to-be-retrieved object had been previously located during encoding. The other three quadrants were semi-randomly labeled ``first'', ``second,'' third'' (e.g., when the critical quadrant was in the top left, the ``first'' quadrant was the top right quadrant, but when the critical quadrant was in the top right, ``first'' corresponded to bottom right, etc.). In both the fixed- and free-viewing condition, participants directed a larger proportion of looks to the critical quadrant (see Figure \ref{fig:E2-gaze-fig-both-conds}). This bias appeared larger in the free-viewing condition, suggesting that the manipulation was (somewhat) effective.

@@

The proportions of looks across quadrants in the free-viewing condition were analyzed using a linear mixed-effects model with quadrant as the predictor (critical as the reference level). The model included random intercepts and slopes for participants\footnote{ \texttt{lme4} syntax: \texttt{lmer(proportion\ \textasciitilde{}\ quadrant\ +\ (1+quadrant\textbar{}subject\_id))}. Among other limitations, this approach violates the independence assumptions of the linear model because looks to the four locations are not independent. This analysis was chosen because it is analogous to the ANOVA analysis conducted in the original paper.}. Proportions of looks were significantly higher for the critical quadrant compared to the other three (first: \emph{b} = -0.06, \emph{SE} = 0.01, \emph{p}\textless0.001, second: \emph{b} = -0.08, \emph{SE} = 0.01, \emph{p}\textless0.001, third: \emph{b} = -0.05, \emph{SE} = 0.01, \emph{p}\textless0.001 )

@@

>>>>

\abstract{%
Web-based data collection allows researchers to recruit large and diverse samples with fewer resources than lab-based studies require. Recent innovations have expanded the set of methodolgies that are possible online, but ongoing work is needed to test the suitability of web-based tools for various research paradigms. Here, we focus on webcam-based eye-tracking; we tested whether the results of five different eye-tracking experiments in the cognitive psychology literature would replicate in a webcam-based format. Specifically, we carried out five experiments by integrating two javascript-based tools: js.psych and a modified version of Webgazer.js. In order to represent a wide range of applications of eye-tracking to cognitive psychology, we chose two psycholinguistic experiments, two memory experiments, and a decision-making experiment. These studies also varied in the type of eye-tracking display, including screens split into halves (Exps. 3 and 5) or quadrants (Exps. 2 and 4), or composed scenes with regions of interest that varied in size (Exp. 1). Outcomes were mixed. The least successful replication attempt was Exp. 1; we did not obtain a condition effect in our remote sample (1a), nor in an in-lab follow-up (1b). However, the other four experiments were more successful, replicating a blank-screen effect (Exp. 2), a novelty preference (Exp. 3), a verb bias effect (Exp. 4), and a gaze-bias effect in decision-making (Exp. 5). These results suggest that webcam-based eye tracking can be used to detect a variety of cognitive phenomena, including those with sensitive time, although paradigms that require high spatial resolution should be adapted to coarser quadrant or split-half displays.
}

@@

During the calibration, dots appeared on the screen one at a time in
different locations, and the participants had to fixate them and click
on each one. Once they clicked on a dot, it would disappear and a new
one would appear in a different location on the screen. The locations of
calibration dots were specific to each experiment (details below) and
appeared in the areas of the screen where the visual stimuli would
appear during the main task in order to ensure that eye movements were
accurately recorded in the relevant regions of interest. After the
calibration was completed, the validation began. Participants were asked
to go through the same steps as the calibration, except that they only
fixated the dots as they appeared in different locations on the screen.
If accuracy on the validation was too low (fewer than 50\% of looks
landed within a 200 px radius of the validation points),
participants were given an opportunity to re-start the calibration and
validation steps.

@@

\subsection{Data Pre-processing}\label{data-pre-processing}

We used R (Version 4.2.1; R Core Team, 2021) and the R-packages \emph{afex} (Version 1.3.0; Singmann, Bolker, Westfall, Aust, \& Ben-Shachar, 2021), \emph{broom.mixed} (Version 0.2.9.4; Bolker \& Robinson, 2020), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, \& Müller, 2021), \emph{forcats} (Version 1.0.0; Wickham, 2021a), \emph{ggplot2} (Version 3.4.4; Wickham, 2016), \emph{jsonlite} (Version 1.8.8; Ooms, 2014), \emph{lme4} (Version 1.1.35.1; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lmerTest} (Version 3.1.3; Kuznetsova, Brockhoff, \& Christensen, 2017), \emph{Matrix} (Version 1.6.5; Bates \& Maechler, 2021), \emph{papaja} (Version 0.1.2; Aust \& Barth, 2020), \emph{readr} (Version 2.1.5; Wickham \& Hester, 2020), \emph{shiny} (Version 1.8.0; Chang et al., 2021), \emph{stringr} (Version 1.5.1; Wickham, 2019), \emph{tidyr} (Version 1.3.1; Wickham, 2021b), and \emph{tinylabels} (Version 0.2.4; Barth, 2022) for all our analyses.

\section{Experiment 1a}\label{experiment-1a}

@@

@@

@@

\subsubsection{Materials and Design}\label{materials-and-design}

Each participant was presented with 16 critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
16 fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.

@@

@@

@@

@@

For each sentence, the target time window began at the onset of the verb and ended 2000 milliseconds later. This window was then divided into 50-ms bins; for each participant and each trial, we recorded whether each object was fixated during the 50-ms bin. Collapsing over trials and participants, and averaging across distractors, we calculated the cumulative probability of fixation, shown in Figure \ref{fig:E1a-spaghetti-fig}, Panel (b).

@@

In our first two analyses, we asked whether participants looked more to the target than to the distractor during the predictive time window, given that the verb is restricting. The first model tested whether there were more fixations to the target object than to the distractor in the time window before the onset of the target noun. We ran a regression model predicting the cumulative fixation probability in the last 50-ms bin before noun onset from the verb condition (restricting = 1 vs.~non-restricting = 0), object type (target = 1 vs.~distractor = 0), and their interaction, along with random effects for participants and images (with no covariance between random effects because the model cannot converge with full covariance matrix)\^{}{[} \texttt{lme4} syntax: \texttt{lmer\_alt(probability\ \textasciitilde{}\ object\_type*verb\_condition\ +\ (object\_type*verb\_condition\ \textbar{}\textbar{}\ subject)\ +\ (object\_type*verb\_condition\ \textbar{}\textbar{}\ scene)}. There were no significant effects, although the critical interaction was in the expected direction (\emph{b} = 0.05, \emph{SE} = 0.03, \emph{p}=0.15).

@@

Altmann and Kamide tested a second model, aligning the predictive time window with the offset of the verb rather than the onset of the noun as above. When we did the same\^{}{[} \texttt{lme4} syntax: \texttt{lmer\_alt(probability\ \textasciitilde{}\ object\_type*verb\_condition\ +\ (object\_type*verb\_condition\ \textbar{}\textbar{}\ subject)\ +\ (object\_type*verb\_condition\ \textbar{}\textbar{}\ scene)}, we again saw that the critical interaction is not significant but numerically in the expected direction (\emph{b} = 0.05, \emph{SE} = 0.03, \emph{p}=0.20).

@@

Finally, we addressed whether participants looked to the target faster in the restrictive vs.~the non-restrictive condition, starting after the onset of the verb. On average, participants looked to the target 335 ms after the noun onset in the restrictive condition James, Minnihan, \& Watson (2023) and 435 ms after the noun onset in the non-restrictive condition James et al. (2023). Thus, first fixations were not only delayed relative to those in the previous studies compared here, but also showed a smaller difference between conditions.

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes\^{}{[} \texttt{lme4} syntax: \texttt{lmer\_alt(time\ \textasciitilde{}\ verb\_condition*verb\_duration\ +\ (verb\_condition\ \textbar{}\textbar{}\ subject)\ +\ (verb\_condition\ \textbar{}\textbar{}\ scene)}. There were no significant effects; participants looked sooner at the target in the restrictive condition, while accounting for verb duration and its interaction with condition, but this was not a statistically significant effect (\emph{b} = -121.60, \emph{SE} = 85.43, \emph{p}=0.17).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1-calib-effect-1.pdf}
\caption{\label{fig:E1-calib-effect}Calibration scores plotted against the verb effect on first fixations (the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition).}
\end{figure}

\subsubsection{Calibration}\label{calibration}

@@

We tested whether a participant's calibration quality was correlated with their effect size. There were three effects of interest: the verb-by-object interaction in predicting fixation probabilities, both in the (1) pre-noun-onset and (2) pre-verb-offset windows (calculated as the difference in target-over-distractor preference between verb conditions), and (3) the effect of verb on the timing of the first target fixation (calculated as the difference in target latency between verb conditions). Across the three effects of interest, calibration quality was not significantly correlated (Effect 1: Pearson's \emph{r} = 0.03, \emph{p} = 0.83, Effect 2: Pearson's \emph{r} = -0.05, \emph{p} = 0.73, Effect 3: Pearson's \emph{r} = 0.10, \emph{p} = 0.49. Figure~\ref{fig:E1-calib-effect} plots the relation between calibration scores and Effect 3. The figure makes clear that there is a large proportion of participants with calibration scores under 50\%.

\paragraph{Re-analysis After Exclusions}\label{re-analysis-after-exclusions}

We re-analyzed the data after excluding participants with calibration scores under 50\% (N = 35). The following analyses thus included 22 participants.

We tested the same three models under these more aggressive exclusion criteria. The first two models, comparing target and distractor fixations in the predictive window, produced very similar results; the critical interaction was not statistically significant (Pre-noun-onset window: \emph{b} = 0.07, \emph{SE} = 0.06, \emph{p}=0.23; Pre-verb-offset window: \emph{b} = 0.05, \emph{SE} = 0.05, \emph{p}=0.28). However, the final model, which tested the effect of verb condition on the timing of fixations to the target, yielded a statistically significant result, unlike in the previous set of analyses with all participants included (\emph{b} = -193.35, \emph{SE} = 96.33, \emph{p}=0.05).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1a-spaghetti-fig-1.pdf}
\caption{\label{fig:E1a-spaghetti-fig}Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.}
\end{figure}

Across three different tests of the hypothesis that listeners will use verb semantics to anticipate the upcoming referent, we found results that were numerically consistent with the hypothesis but not statistically significant. A comparison to published data (James et al., 2023) demonstrates that looks to the objects in the scene (relative to background or off-screen looks) were depressed across conditions and objects. After eliminating participants with validation accuracy under 50\% and/or 10\% or fewer fixations to any ROIs, we were left with only 22 of the original 60 participants. Analyses on this subset resulted in one effect reaching statistical significance (shorter target fixation latencies in the restricting vs.~non-restricting verb condition).
Given that nearly two-thirds of participants had poor data quality, we ran a follow-up webcam study in a lab setting order to isolate the potential causes.

\section{Experiment 1b}\label{experiment-1b}

Experiment 1b tested whether the failure to replicate significant condition effects in Experiment 1a was due to features of conducting the study remotely (i.e.~varied experimental settings and apparatuses, lower compliance) rather than webcam-based eye-tracking or \texttt{webgazer} \emph{per se}. Thus, Experiment 1b took place in a lab setting with undergraduate participants but otherwise used the same Method as Experiment 1a.

Forty-nine participants completed the study in a lab setting. They were recruited via the Boston College subject pool. Participants needed to be 18 years of age or older and native speakers of English.

\subsubsection{Materials and Design}\label{materials-and-design-1}

Materials were identical to those in Experiment 1a.

After being greeted by the experimenter and completing the informed consent form, participants followed on-screen prompts to complete the study, including calibration, as described in the Experiment 1a Procedure.

\subsubsection{Apparatus}\label{apparatus}

HELP

@@

For each sentence, the target time window began at the onset of the verb and ended 2000 milliseconds later. This window was then divided into 50-ms bins; for each participant and each trial, we recorded whether each object was fixated during the 50-ms bin. Collapsing over trials and participants, and averaging across distractors, we calculated the cumulative probability of fixation, shown in Figure \ref{fig:E1b-spaghetti-fig}, Panel (b). The results from Experiment 1a are copied here for ease of comparison (Panel a).

@@

Finally, we addressed whether participants looked to the target faster in the restrictive vs.~the non-restrictive condition, starting after the onset of the verb. On average, participants looked to the target 405 ms after the noun onset in the restrictive condition and 399 ms after the noun onset in the non-restrictive condition. As in the remote sample, the latencies are overall slower than in results published by Altmann \& Kamide (1999) and James et al.~(2023). Unlike in the remote sample, however, the difference is in the unexpected direction, such that participants looked to the target faster in the non-restrictive condition.

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes. There were no significant effects; results revealed that the paradoxical advantage in the non-restrictive condition was not statistically significant (\emph{b} = 2.88, \emph{SE} = 117.68, \emph{p}=0.98). Effects of verb duration and its interaction with condition were also not statistically significant (duration: \emph{b} = -0.58, \emph{SE} = 0.54, \emph{p}=0.30; interaction: \emph{b} = -0.60, \emph{SE} = 0.96, \emph{p}=0.53 )

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1b-calib-effect-1.pdf}
\caption{\label{fig:E1b-calib-effect}Calibration scores plotted against the verb effect, operationalized as the latency of the first target fixation in the non-restricting verb condition minus that in the restricting verb condition.}
\end{figure}

\subsubsection{Calibration}\label{calibration-1}

As before, participants' calibration quality was measured as the mean percentage of fixations that landed within 200 pixels of the calibration point. Calibration quality ranged from 5.13\% to 97.89\%.
Across the three condition effects of interest, calibration quality was not significantly correlated (pre-noun: Pearson's \emph{r} = -0.24, \emph{p} = 0.11; pre-verb-offset: Pearson's \emph{r} = -0.19, \emph{p} = 0.20; first fixation: Pearson's \emph{r} = -0.16, \emph{p} = 0.29. Figure~\ref{fig:E1-calib-effect} plots the relation between calibration scores and Effect 3.

\paragraph{Re-analysis After Exclusions}\label{re-analysis-after-exclusions-1}

Again, we excluded participants with validation accuracy under 50\%, which eliminated 23 participants. The following analyses included 26 participants.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1b-spaghetti-fig-1.pdf}
\caption{\label{fig:E1b-spaghetti-fig}Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.}
\end{figure}

Overall, the results of Experiment 1 paint a sobering picture of web-based eye-tracking. In Experiment 1a, results from the remote samplewere in the expected direction but effects were smaller and delayed relative to previous work and failed to reach statistical significance. To test whether this could be explained by the variability in experimental settings across participants, we replicated our procedure in a lab setting. Surprisingly, the results in the in-lab study were less aligned with previous work; critical effects were not significant nor were they in the expected direction numerically. This suggests the problem was not with running the study online or due to working with a more diverse population via Prolific.

Taken together, the instability of these effects might suggest that this paradigm is not well-suited for webcam-based eye-tracking. Notably, the ROIs were tightly drawn around the five to six objects in each scene (drawing larger ROIs in these scenes would have led to overlapping objects) and thus, analyses were unforgiving of inaccurate calibration. Further evidence comes from a recent \texttt{webgazer} study that successfully replicated a modified version of the Altmann and Kamide study with only four objects, each in separate quadrants, allowing for larger, more distinct ROIs (Prystauka et al., 2023).

@@

Sixty participants were paid for their participation.
The sample size was motivated in part by budget constraints, but was
nonetheless 2.5x larger than the original sample size of 24).
Data from one participant were not properly
recorded due to unknown technical issues, so data from 59 participants
were included in all analyses to follow.

\subsubsection{Materials and Design}\label{materials-and-design-2}

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. The two blocks differed by whether the recall phase was in the free-viewing or fixed-viewing condition, as described in the Procedure. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first.
During each encoding phase, participants saw a
grid indicating the four quadrants of the screen. Each quadrant
contained six cartoon images of items belonging to the same category. The four categories were humanoids, household objects, animals, and
methods of transportation (see Figure~\ref{fig:E2-example-trial}). Different images were used in each block; there were 48 unique images total across the experiment.

Each recall phase presented participants with a blank screen with a central fixation cross as they listened to true/false statements testing their recall of the previous grid. Each statement fell into either an interobject or intraobject
condition. Interobject statements were those that compared two different
items in the grid (e.g.~``The skeleton is to the left of the robot''),
while intraobject statements were those that asked about the orientation
of a single item (e.g.~``The bus is facing right''). There were 48 total
statements in each block, such that each object was the subject of both an intraobject and an intraobject statement; there were 96 unique statements total across the experiment.

@@

Participants received instructions that included an example grid and an explicit request that they not use any tools to help them during encoding. Participants then began their first encoding phase. Each of the four quadrants was presented one at a time. First, a list of
the items in the quadrant was shown, then the pictures of items were displayed in the quadrant.
For each item, participants used their arrow keys to indicate whether the object was facing left or right. After the participant identified the direction of each
item, they would have an additional 30 seconds to encode the name and
orientation of each item in the quadrant. Finally, after all four quadrants
were presented, participants were shown the full grid of
24 items and had 60 seconds to further encode the name and orientation
of each item.

Participants then entered the first recall phase, in which they listened to the 48 statements and
responded by pressing the `F' key for false statements and `T' for true
ones. While listening to these statements, in
the free-viewing block, participants saw a blank screen and were allowed
to freely gaze around the screen. During the fixed-viewing block,
participants were asked to fixate a small cross in the center of the
screen throughout the recall phase. In both cases, the mouse was
obscured from the screen.

Participants then proceded to the second encoding and recall phases as described above. After completing both encoding-recall blocks, participants were asked to
answer a few survey questions (such as whether they wore glasses or
encountered any distractions).

@@

\subsubsection{Eye-gaze}\label{eye-gaze}

Looks during the retrieval period were categorized as belonging to one of four quadrants based on the x,y coordinates. The critical quadrant was the one in which the to-be-retrieved object had been previously located during encoding. The other three quadrants were labeled ``first'', ``second,'' and ``third'' depending upon the location of the critical quadrant (e.g., when the critical quadrant was in the top left, the ``first'' quadrant was the top right quadrant, but when the critical quadrant was in the top right, ``first'' corresponded to bottom right, etc.). In both the fixed- and free-viewing condition, participants directed a larger proportion of looks to the critical quadrant (see Figure \ref{fig:E2-gaze-fig-both-conds}). This bias appeared larger in the free-viewing condition, suggesting that the manipulation was somewhat effective.

@@

The proportions of looks across quadrants in the free-viewing condition were analyzed using a linear mixed-effects model with quadrant as the predictor (critical as the reference level). The model included random intercepts and slopes for participants\footnote{ \texttt{lme4} syntax: \texttt{lmer(proportion\ \textasciitilde{}\ quadrant\ +\ (1+quadrant\textbar{}subject\_id))}. Among other limitations, this approach violates the independence assumptions of the linear model because looks to the four locations are not independent. This analysis was chosen because it is analogous to the ANOVA analysis conducted in the original paper.}. Proportions of looks were significantly higher for the critical quadrant compared to the other three (first: \emph{b} = -0.06, \emph{SE} = 0.01, \emph{p}\textless0.001, second: \emph{b} = -0.08, \emph{SE} = 0.01, \emph{p}\textless0.001, third: \emph{b} = -0.05, \emph{SE} = 0.01, \emph{p}\textless0.001).

@@

\subsubsection{Materials and Design}\label{materials-and-design-3}
